---
title: "Gov 50 Final Project"
author: "Jake Brownley"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## INTRODUCTION

  The National Basketball Association (NBA) is one of the most polarizing professional sports leagues in the world. Competing with the likes of professional football, professional baseball, and international soccer, the NBA is one of the most viewed sports among Americans. In recent years, the NBA has seen an increase in "load-management" - the idea that, in order to succeed, the best players on the best teams need to rest periodically throughout the regular season to be fit to win a championship in the postseason. Unsurprisingly, the NBA has received negative feedback regarding load-management. The main argument against the idea is that fans pay their hard-earned money to come watch the best, most celebrated players play. 
  
  Load management has sparked intense conversations in the world of sports, with "experts" saying that it seems like the best players actually play the least. This leads me to my research question: Do higher-paid NBA players play fewer games per year than players paid less than them? I, like the so called experts, agree that it does appear that the highest-paid players actually play the least. Therefore, I hypothesize that players who are paid the most tend to play fewer games per year than a player paid less. In my findings, I will help find the real truth behind the load-management dispute. 

## DATA SECTION

  The data source being evaluated is the NBA 2022-2023 salaries dataset referred to as `nba_salaries`. This dataset contains statistics for all 466 players from the 2022-2023 season, including information on salary, games played, minutes played, field goal average, and field goal percentage, among others. To address my research question, I will focus specifically on salary and games played. 
  
  In this context, the dependent variable is `games played` as it represents the outcome that I am trying to understand and predict. Conversely, the independent variable is `salary` as it is the factor that I believe may have an effect on the amount of games played (the dependent variable). Both variables are measured through statistical observation. Simply put, salary is determined by the player's earnings in the 2022-2023 season, while games played are recorded based on the number of games a player participated in during the same season. The plot below illustrates the dependent variable.
  
  The research design employed is cross-sectional. A cross-sectional study is an observational research design used to compare outcomes between treated and control units at a single point in time. This dataset constitutes an observational study involving the collection of data from all NBA players during the 2022-2023 season.
  

```{r}
library(tidyverse)
library(ggplot2)

nba_salaries <- read.csv("nba_salaries.csv")

```

```{r}
library(tidyverse)
library(ggplot2)

nba_dependent_variable <- 
  ggplot(data = nba_salaries, aes(x = GP, y = )) +
  geom_histogram(binwidth = 1, fill = "red", color = "black") +
  labs(title = "Games Played Distribution", x = "Games Played (GP)", y = "Amount of Players")
nba_dependent_variable
```


## RESULTS SECTION

  Below, the relationship between salary and games played is depicted. High density is observed around both low pay and fewer games played, suggesting that, according to the 'Games Played vs. Salary' plot, individuals who are paid more indeed play more than their counterparts. However, to determine the true proportion of games played and salary, regression analysis is necessary.
  
  The primary coefficient of interest in this regression is salary. The objective is to identify the correlation between games played and salary. In the analysis, it is apparent that every dollar paid is associated with 0 games, but the actual value is not 0. Technically, one dollar paid corresponds to 7.916e-07 games played. Consequently, the p-value for 'Dollars Per Game' is less than 0.001, indicating statistical significance. Moreover, with the statistically significant data, it can reasonably be concluded that the null hypothesis is rejected. This coefficient represents a causal effect— the amount of money paid has a direct impact on a player's participation. The positive regression indicates that the results are causally driven.
  
  Following the regression evaluation, a table displays the number of games played for players ranked 1-100 based on pay and players ranked 101-350 based on pay. To simplify the results from the regression table, the columns 'Top_100_Games_Played_Avg' and 'Top_101to350_Games_Played_Avg' illustrate the average number of games played in the 2022-2023 NBA season for different pay groups. Notably, players paid in the top 100 played almost 7 more games than those ranked 101-350 based on pay.
  
## Plot of Main Analysis

```{r}
library(plotly)

results_scatterplot <- 
  ggplot(data = nba_salaries, aes(x = Salary, y = GP)) +
  geom_point(color = "blue") + 
  labs(title = "Games Played vs. Salary", x ="Salary", y = "Games Played (GP)")

ggplotly(results_scatterplot)
```
## Main Regression of Interest

```{r}
library(tidyverse)
library(broom)

fit <- lm(nba_salaries$GP ~ nba_salaries$Salary)

var_labels <- c(
  "(Intercept)" = "Intercept",
  "nba_salaries$Salary" = "Dollars Per Game")

modelsummary::modelsummary(list(fit),
                           coef_map = var_labels,
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))

```
## Games Played by Specific Salary Ranges

```{r}
gamesplayed_100 <- read_csv("nba_salaries.csv") |>
  slice_max(GP, n = 100) |>
  summarize(Top_100_Games_Played_Avg = mean(GP))

gamesplayed_101to350 <- read_csv("nba_salaries.csv") |>
  slice_max(GP, n = 101-350) |>
  summarize(Top_101to350_Games_Played_Avg = mean(GP))

SalaryandGP_comparison_tibble <- tibble(gamesplayed_100, 
                                  gamesplayed_101to350)
SalaryandGP_comparison_tibble
knitr::kable(SalaryandGP_comparison_tibble)
```



## CONCLUSION

As mentioned earlier, the value of 7.916e-07 (0.0000007916) may seem insignificant on its own. However, in the context of NBA salaries, where payments are in the millions, the significance becomes evident. When properly scaled, it is determined that for every million dollars paid, a player participates in 0.7916 games—almost a whole game. This finding is undoubtedly statistically significant, establishing a clear correlation between higher pay and increased playing time. In the table above, it's evident that players paid above the league average played seven more games in the 2022-2023 NBA season than their counterparts. This result contradicts my initial hypothesis: players who are paid the most tend to play fewer games per year than those paid less. Unfortunately, the NBA experiences a yearly portion of players missing games due to injuries. In my research, injuries pose a threat to inference. Players participating in only a handful of games due to injuries skew the data. Addressing this issue is crucial. While removing injured players would yield a more accurate dataset, it would also be unnatural as injuries are an inherent part of the game. To enhance this analysis further, a more in-depth examination of salary is warranted. I compared games played between the top two pay groups. To uncover additional insights, exploring another salary group—specifically, those paid the least—would be valuable. Rookies and free agents often receive minimum pay, making it intriguing to assess their games played production relative to their salary